---
title: "Strand AI Platform Quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Strand AI Platform Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette walks end-to-end through the [Strand AI Platform](https://strandai.com)
API using the `strand` R package: upload a Whole Slide Image (WSI), submit a
Lattice H&E to multiplex protein inference job, wait for completion, and pull
results back as a `SpatialExperiment`.

> The chunks below set `eval = FALSE` so the vignette renders cleanly on CRAN
> without network access. To run them yourself, mint an API key at
> `app.strandai.com/settings/api-keys`, set `STRAND_API_KEY`, and run
> interactively.

## 1. Set up

```{r}
library(strandai)

# Reads STRAND_API_KEY; for local development:
# Sys.setenv(STRAND_BASE_URL = "http://localhost:3000")
client <- strand_client()
client
```

## 2. Upload a slide

`strand_upload_file()` initiates a resumable session and streams the file to
GCS in 8 MiB chunks. The final PUT triggers automatic ingest, and the function
polls only until the upload status leaves `uploading`. Dimensions can remain
unavailable while preprocessing continues; estimation reports when the sample
is ready.

```{r}
upload <- strand_upload_file(client, "slide.svs", progress = TRUE)
upload
```

## 3. Work with samples

`strand_samples_list()` lists owned samples, the public cohort, or both.
Every item has an `ownership` discriminator and a canonical `id` accepted by
`strand_samples_get()`. Owned detail includes recent inference jobs; public
detail contains public geometry and marker channels without job history.

```{r}
page <- strand_samples_list(client, scope = "all")
sample <- strand_samples_get(client, page$items[[1]]$id)

if (sample$ownership == "mine") {
  sample <- strand_patch_sample(
    client, sample$id,
    name = "Baseline biopsy",
    tags = c("baseline", "responding"),
    mpp = 0.26
  )
}
```

## 4. Price credits

A dry run performs the same validation and pricing as submission without
reserving credits or creating a job.

```{r}
markers <- c("CD3", "CD8", "Ki67", "PD-L1")
est <- strand_predict(client, upload$id, markers, dry_run = TRUE)
message(
  est$patch_count, " patches x ", est$marker_count, " markers = ",
  est$estimated_credits, " credits"
)
message("org balance: ", est$org_balance)
```

## 5. Submit and wait

Submission **atomically reserves** the estimated credits in the same database
transaction that creates the job. If your org has insufficient credits, you'll
see a `strand_insufficient_credits_error`.

```{r}
job <- strand_predict(client, upload$id, markers)
status <- strand_job_wait(job, progress = TRUE)
stopifnot(status$status == "completed")
```

### Choosing a model

Pass `model = "v0.7"` to select the current dispatchable Lattice version
explicitly. Omitting `model` lets the platform pick the same current default.
Older `"v0.4"` and `"v0.5"` labels can appear on historical jobs but are
sunset for new submissions.

```{r}
job <- strand_predict(client, upload$id, markers, model = "v0.7")
```

### Fire-and-forget

For long pipeline runs, pass `wait = FALSE` to `strand_run()` to return the
job handle as soon as upload + submit complete. The job continues server-side
— you can poll or download later from the same or a fresh process.

```{r}
job <- strand_run(client, "slide.svs", markers, wait = FALSE)
# ...later:
status <- strand_job_wait(job)
spe <- strand_download_results(job)
```

## 6. Download as SpatialExperiment

`strand_download_results()` walks the OME-Zarr result store via the
API-key-authenticated proxy and assembles the channels into a
[`SpatialExperiment`](https://bioconductor.org/packages/SpatialExperiment).
Rows are markers, columns are pixels; spatial coordinates are stored in
`spatialCoords(spe)`.

For very large slides, supply `path = "out/"` to mirror the zarr store to disk
and read selectively with `Rarr` instead.

```{r}
spe <- strand_download_results(job)
spe
```

## Error handling

```{r}
tryCatch(
  strand_predict(client, upload$id, markers),
  strand_insufficient_credits_error = function(e) {
    message("Need ", e$required, " more credits.")
  }
)
```
