---
title: "Strand Platform Quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Strand Platform Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette walks end-to-end through the [Strand Platform](https://strandai.com)
API using the `strand` R package: upload a Whole Slide Image (WSI), submit a
Lattice H&E to multiplex protein inference job, wait for completion, and pull
results back as a `SpatialExperiment`.

> The chunks below set `eval = FALSE` so the vignette renders cleanly on CRAN
> without network access. To run them yourself, mint an API key at
> `app.strandai.com/settings/api-keys`, set `STRAND_API_KEY`, and run
> interactively.

## 1. Set up

```{r}
library(strandai)

# Reads STRAND_API_KEY; for local development:
# Sys.setenv(STRAND_BASE_URL = "http://localhost:3000")
client <- strand_client()
client
```

## 2. Upload a slide

`strand_upload_file()` handles the resumable upload flow end-to-end: it
initiates a session, streams the file to GCS in 8 MiB chunks, and finalizes
once all bytes are written. Slide dimensions are read on finalize and used to
estimate credits.

```{r}
upload <- strand_upload_file(client, "slide.svs", progress = TRUE)
upload
```

## 3. Estimate credits

Estimation is free — call it as many times as you like while iterating on a
marker panel.

```{r}
markers <- c("CD3", "CD8", "Ki67", "PD-L1")
est <- strand_estimate(client, upload$id, markers)
message(
  est$patch_count, " patches x ", est$marker_count, " markers = ",
  est$estimated_credits, " credits"
)
message("org balance: ", est$org_balance)
```

## 4. Submit and wait

Submission **atomically reserves** the estimated credits in the same database
transaction that creates the job. If your org has insufficient credits, you'll
see a `strand_insufficient_credits_error`.

```{r}
job <- strand_predict(client, upload$id, markers)
status <- strand_job_wait(job, progress = TRUE)
stopifnot(status$status == "completed")
```

### Choosing a model

Pass `model = "v10"` for the original 7-marker panel or
`model = "v10-fullpanel"` for the 192-marker sibling. Both share GenePT
embeddings, so the marker vocabulary is identical between them — picking a
model is a weights swap, not a vocab swap. Omitting `model` lets the platform
pick (currently `v10-fullpanel`).

```{r}
job <- strand_predict(client, upload$id, markers, model = "v10-fullpanel")
```

### Fire-and-forget

For long pipeline runs, pass `wait = FALSE` to `strand_run()` to return the
job handle as soon as upload + submit complete. The job continues server-side
— you can poll or download later from the same or a fresh process.

```{r}
job <- strand_run(client, "slide.svs", markers, wait = FALSE)
# ...later:
status <- strand_job_wait(job)
spe <- strand_download_results(job)
```

## 5. Download as SpatialExperiment

`strand_download_results()` walks the OME-Zarr result store via the
API-key-authenticated proxy and assembles the channels into a
[`SpatialExperiment`](https://bioconductor.org/packages/SpatialExperiment).
Rows are markers, columns are pixels; spatial coordinates are stored in
`spatialCoords(spe)`.

For very large slides, supply `path = "out/"` to mirror the zarr store to disk
and read selectively with `Rarr` instead.

```{r}
spe <- strand_download_results(job)
spe
```

## Error handling

```{r}
tryCatch(
  strand_predict(client, upload$id, markers),
  strand_insufficient_credits_error = function(e) {
    message("Need ", e$required, " more credits.")
  },
  strand_rate_limit_error = function(e) {
    message("Org over the concurrent-job cap; retry in ", e$retry_after, "s.")
  }
)
```
